--- 
title: "Women's and Minorities' Working Opportunities in New York City from 1850 to 1940"
author: "Luwei Zhang, Guangyu Wu, Jingxiang Zhang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

At the turn of the 20th century, New York City experienced one of the fastest periods of urban growth, as it overtook London to become the most populous urbanized area in the world and the first megacity in the human history^["New York Urbanized Area: Population & Density from 1800 (Provisional)". Demographia. Retrieved July 8, 2009.]. Powering this tremendous growth is the working people of NYC, who came from different backgrounds and contributed to the booming industries of the city. 

However, working opportunities are not always abundant. The fierce competition between disadvantaged groups such as the Irish immigrants and the Black population escalates to one of the worst riots in American history. 

In the meantime, the working conditions of women 

under-regulated industry "the Triangle Shirtwaist Factory fire, the city's worst industrial disaster, took the lives of 146 garment workers and spurred the growth of the International Ladies' Garment Workers' Union and major improvements in factory safety standards."^["The 1911 Triangle Factory Fire". Kheel Center, Cornell University. Retrieved February 9, 2017.] 

<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
knitr::write_bib("knitr", file = "packages.bib")
## Tidyverse
install.packages("dplyr")
install.packages("ggplot2")
install.packages("stringr")
install.packages("purrr")
## HTML tables
install.packages("DT")
## GIS
install.packages("sf")
```

## IPUMS introduction

Data ^[Steven Ruggles, Sarah Flood, Sophia Foster, Ronald Goeken, Jose Pacas, Megan Schouweiler and Matthew Sobek. IPUMS USA: Version 11.0 [dataset]. Minneapolis, MN: IPUMS, 2021. 
https://doi.org/10.18128/D010.V11.0] 

## Scope

## Creating Data Extract

## IPUMSR

```{r}
library(ipumsr) 
library(tidyverse) 
library(sf)
```

```{r}
ddi <- read_ipums_ddi("./data/usa_00038.xml")
data <- read_ipums_micro(ddi)
data <- data %>% select(-c(SERIAL, PERNUM, RELATED, RACED, BPLD, MBPLD, FBPLD, YRSUSA2, MTONGUE, MTONGUED, PRESGL))
print(names(data))
```

## Codebook 


<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

## Decoding, Handling n/a, missing, and other category codes, Enforce data type, Reduce cardinality

```{r}
library(ipumsr) 
library(tidyverse) 
library(sf)
```

```{r}

ddi <- read_ipums_ddi("./data/usa_00038.xml")
data <- read_ipums_micro(ddi)
data <- data %>% select(-c(SERIAL, PERNUM, RELATED, RACED, BPLD, MBPLD, FBPLD, YRSUSA2, MTONGUE, MTONGUED, PRESGL))

# OCC1950CAT, IND1950CAT (augmentation)
occ_cat <- read_csv("./data/occupation_categories.csv")
ind_cat <- read_csv("./data/industry_categories.csv")
data <- left_join(data, occ_cat, by ='OCC1950')
data <- left_join(data, ind_cat, by ='IND1950')

# NA relabeling
data$OCC1950 <- lbl_na_if(data$OCC1950, ~.val > 970) %>% as_factor()
data$IND1950 <- lbl_na_if(data$IND1950, ~.val >= 975 | .val == 0) %>% as_factor()
data$OCCSCORE <- as.numeric(data$OCCSCORE)

df <- as_factor(data)
# YEAR
df$YEAR <- as.numeric(as.character(df$YEAR))
# CITY 
## table(df$CITY)
# FAMSIZE
df$FAMSIZE <- as.character(df$FAMSIZE)
df$FAMSIZE[df$FAMSIZE == "1 family member present"] = "1"
df$FAMSIZE[df$FAMSIZE == "2 family members present"] = "2"
df$FAMSIZE <- as.numeric(df$FAMSIZE)
# RELATE
df$RELATE <- as.character(df$RELATE)
df$RELATE[df$RELATE == "Head/Householder"] = "Head"
df$RELATE[df$RELATE == "Child-in-law"] = "Child"
df$RELATE[df$RELATE == "Grandchild"] = "Child"
df$RELATE[df$RELATE == "Parent-in-Law"] = "Parent"
df$RELATE[df$RELATE == "Sibling-in-Law"] = "Sibling"
df$RELATE[df$RELATE == "Other relatives"] = "Relative"
df$RELATE[df$RELATE == "Partner, friend, visitor"] = "Relative"
df$RELATE[df$RELATE == "Other non-relatives"] = "Other"
df$RELATE[df$RELATE == "Institutional inmates"] = "Inmate"
df$RELATE <- as.factor(df$RELATE)
# SEX
## table(df$SEX)
# AGE
df$AGE <- as.character(df$AGE)
df$AGE[df$AGE == "Less than 1 year old"] = "0"
df$AGE[df$AGE == "90 (90+ in 1980 and 1990)"] = "90"
df$AGE[df$AGE %in% c("100 (100+ in 1960-1970)", "112 (112+ in the 1980 internal data)", "115 (115+ in the 1990 internal data)")] = "100"
df$AGE <- as.numeric(df$AGE)
# MARST
df$MARST <- as.character(df$MARST)
df$MARST[df$MARST == 'Married, spouse present'] = 'Married'
df$MARST[df$MARST == 'Never married/single'] = 'Single'
df$MARST[df$MARST %in% c('Married, spouse absent','Separated','Divorced','Widowed')] = 'Widowed or Divorced'
df$MARST <- as.character(df$MARST)
df$MARST <- as.factor(df$MARST)
# RACE
df$RACE <- as.character(df$RACE)
df$RACE[df$RACE == 'Black/African American/Negro'] = 'Black'
df$RACE[!(df$RACE %in% c('White','Black'))] = 'Other'
df$RACE <- as.factor(df$RACE)
# BPL
# MBPL
# FBPL
# NATIVITY
df$NATIVITY <- as.character(df$NATIVITY)
df$NATIVITY[df$NATIVITY %in% c("Native born, and both parents native born","Native born, and father foreign, mother native","Native born, and mother foreign, father native")] = 'Native born with native parent'
df$NATIVITY[df$NATIVITY == 'Native born, and both parents foreign'] = 'Native born with foreign parent'
df$NATIVITY[df$NATIVITY == 'N/A'] = NA
df$NATIVITY <- as.factor(df$NATIVITY)
# CITIZEN
df$CITIZEN <- as.character(df$CITIZEN)
df$CITIZEN[df$CITIZEN == 'Not a citizen, but has received first papers'] = 'Applying citizenship'
df$CITIZEN[!(df$CITIZEN %in% c('Naturalized citizen','Not a citizen', 'Applying citizenship'))] = NA
df$CITIZEN <- as.factor(df$CITIZEN)
# YRIMMIG
df$YRIMMIG <- as.character(df$YRIMMIG)
df$YRIMMIG = str_replace(df$YRIMMIG, " \\s*\\([^\\)]+\\)", "")
df$YRIMMIG[df$YRIMMIG %in% c('N/A', 'Not reported')] = NA
df$YRIMMIG <- as.numeric(df$YRIMMIG)
# SCHOOL
df$SCHOOL <- as.character(df$SCHOOL)
df$SCHOOL[df$SCHOOL %in% c("N/A", "Unknown", "Missing")] = NA
df$SCHOOL <- as.factor(df$SCHOOL)
# LIT
df$LIT <- as.character(df$LIT)
df$LIT[df$LIT %in% c("Yes, literate (reads and writes)","Cannot read, can write","Cannot write, can read")] = 'Yes, literate'
df$LIT[df$LIT == 'No, illiterate (cannot read or write)'] = 'No, illiterate'
df$LIT[df$LIT %in% c("N/A", "Unknown, illegible or blank")] = NA
df$LIT <- as.factor(df$LIT)
# LABFORCE
df$LABFORCE <- as.character(df$LABFORCE)
df$LABFORCE[df$LABFORCE == 'N/A'] = NA
df$LABFORCE <- as.factor(df$LABFORCE)
# OCC1950
# IND1950
# INCWAGE
df$INCWAGE <- as.character(df$INCWAGE)
df$INCWAGE[df$INCWAGE %in% c('N/A','Missing')] = NA
df$INCWAGE <- as.numeric(df$INCWAGE)
# OCCSCORE
df$OCCSCORE[df$OCCSCORE == 0] = NA
# EDSCOR50
df$EDSCOR50 <- as.character(df$EDSCOR50)
df$EDSCOR50[df$EDSCOR50 %in% c('N/A')] = NA
df$EDSCOR50 <- as.numeric(df$EDSCOR50)

# YRINUSA (augmentation)
df <- df %>%
 mutate(
  YRINUSA = YEAR - YRIMMIG
 )

write.csv(df, file = './data/ipums.csv', row.names = FALSE)

```


<!--chapter:end:03-cleaning.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Missing values

As the one of world's largest collections of open-access census microdata, IPUMS has been meticulously cleaned and analyzed for completeness and missing patterns. According to its documentation, "missing, illegible and inconsistent values" in variables are edited or allocated, and tagged with data quality flags to indicate the kind of edits applied^[Introduction to Data Editing and Allocation. "IPUMS USA." IPUMS.org. Accessed 2021 Nov 18.  https://usa.ipums.org/usa/flags.shtml]. When understanding the missing values in the outcome data, which we will use for this project, we need bear in mind that the above-mentioned sources of missing values are cleaned and interpolated by the IPUMS team to the best they can.

To reduce the visual clutter in the plot of missing patterns, we rename variable names into 3 letter abbreviations and remove the variables that are not related to working conditions such as person unique identifier. In total, there are 306 missing patterns, but tails are almost negligible in terms of their percentage. Thus, we focus on top 30 patterns which made up the majority of the data. To do this, we augment the `plot_missing` function to add a parameter to control how many patterns to display in graph. The results are shown below.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
knitr::write_bib("knitr", file = "packages.bib")

library(tidyverse)
library(patchwork)
source("plot_missing_top.R")
source("plot_missing.R")
```

```{r fig.height = 9, fig.width = 9}

# Check data codebook in help window
ddi <- read_ipums_ddi("./data/usa_00038.xml")
ipums_view(ddi)

# Load data
df <- read.csv(file = './data/ipums.csv',na.strings=c("NA"))

# Abbreviate variable names 
colnames(df) <- c("YEAR", "CITY", "FAMS", "REL", "SEX", "AGE", "MAR", 
"RAC", "BPL", "MBP", "FBP", "NAT", "CIT", "YRI", 
"SCH", "LIT", "LAB", "OCC", "IND", "INC", 
"OCS", "EDS", "OCA", "ICA", "YUS")

# Remove variable irrelevant to working condition analysis
df <- df %>% select (-c("YEAR", "CITY", "FAMS"))

# Plot top 30 missing patterns 
plot_missing_top(df, percent = TRUE, top_patterns = 30)
```

Focusing the proportion of missing values for different variables, we first observe that basic demographic variables, such as `age`, `sex`, `race`, `relationship to head of household`, have no missing value for the entirety of the dataset. This is understandable because these demographic information are what the census bureau primarily want to collect, and these questions are asked consistently across the decades.

Moving to the left side of the top bar chart, we can see that `income`, `year of immigration`, and `citizenship status`, are the three variables with the highest proportion of missing values. The proportion of missing value of `income` is due to the fact that this variable is only introduced in the year 1940, the last of the nine census years that our dataset covers. Similarly, influenced by the evolving US immigration law, `year of immigration` and `citizenship status` fields came in and out of the census survey questionnaires across the century, resulting in higher percentage of missing values. 

The next groups of variables with relatively high level of missing values are all occupation-related. In fact, they almost have the identical proportion of missing values. After digging into the documentation and data transformation process, we quickly realize these variables all stem from two root variable `OCC` (occupation) and `IND` (industry). `OCA` and `ICA` stand for occupation category and industry category, which are higher level categorization of the detailed information. `OCS`, `EDS`, `PRS` are scores that captures the wage level, education level, and prestige level of a certain occupation, thus are tied to the `OCC` variable^[Integrated Occupation And Industry Codes And Occupational Standing Variables In The Ipums. "IPUMS USA." IPUMS.org. Accessed 2021 Nov 18. https://usa.ipums.org/usa/chapter4/chapter4.shtml]. Overall, 39% of the population doesn't have `OCC` information. When inspecting their demographic information, they tend to be younger children or females. 

Now we switch our attention to the missing patterns. One noticeable trend in the patterns is that the occupation related variables tend to be missing together, a phenomenon discussed and explained in the previous paragraph. The most common missing pattern is the population with no occupation information and no year of immigration, which implies they are native-born residents doing non-wage earning house keeping at home. The next three most common patterns are very similar to the first one, and together they account for about one-third of the population. These missing patterns begs the question of why so many people are missing from the workspace and labor force.




<!--chapter:end:04-missing.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

```{r}
library('tidyverse')
```

```{r}
df <- read.csv(file = 'data/ipums.csv')
drop_columns = c('BPL', 'MBPL', 'FBPL', 'NATIVITY', 'CITIZEN', 'YRIMMIG', 
                 'YRINUSA', 'INCWAGE', 'MARST', 'OCC1950', 'IND1950', 'CITY')
data <- df %>%
  select(-one_of(drop_columns)) %>%
  filter(AGE>=18 & AGE<=66)
```

Before analysis, we first do data cleaning and select several features and data for analyzing. The following data contains features including three main categories, which are minority features, education features, and employment features.

1. Minority Feature: Respondents Age, Respondents Race
2. Education Feature: LIT(Whether could read or write or not)
3. Employment Feature: LABFORCE(Whether in the job market), OCCSCORE(Occupational income score), OCC1950CAT(Occupational category in 1950 standard), IND19050CAT(Industry category in 1950 standard)

The following analysis will give a glimpse of the minority working conditions, the influence of education levels, and the working conditions transition across different years.

## Working Conditions

### Gender and Employment

1. Labor force percentage for each gender: From the following figure, we could conclude that most of the male are in the labor force, while half of the female are not in the labor force. There is significant difference in working conditions between male and female.

```{r}
ggplot(data, aes(LABFORCE, group = SEX)) + 
  geom_bar(aes(y = ..prop.., fill = LABFORCE), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  facet_wrap(~SEX) +
  coord_flip() +
  xlab("Labor Force") + 
  ylab("Percentage in Each Gender") +
  ggtitle("Labor Force Percentage for each Gender") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

2. Occupation score for each gender: The occupation score represents the salary for the workers. As shown in the following figure, the occupation score for male workers are significant higher than the female workers.

```{r}
ggplot(data, aes(x=SEX, y=OCCSCORE)) +
  geom_boxplot() +
  coord_flip() +
  ggtitle("Occupation Score for each Gender") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

3. Occupation category for each gender: There is difference between each gender for the occupation categories. Female tends to work as operatives, clerical and kindred and private household service workers. Male tends to work as craftsmen, managersm and laborers.

```{r}
ggplot(data) +
  geom_bar(aes(x=fct_infreq(OCC1950CAT), fill=SEX), stat = "count") +
  xlab("Occupation Category") +
  ylab("Count of Records") + 
  ggtitle("Occupation Category for each Gender") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

4. Industry category for each gender: There is difference between male and female in industry categories. Female tends to work in manufacturing-nondurable goods, personal services, and professional and related services. Male tends to work in durable goods manufacturing, construction, and transportation.

```{r}
ggplot(data) +
  geom_bar(aes(x=fct_infreq(IND1950CAT), fill=SEX), stat = "count") +
  xlab("Industry Category") +
  ylab("Count of Records") + 
  ggtitle("Industry Category for each Gender") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

### Race and Employment

1. Labor force percentage for each Race: As shown in the following figure, there is little difference between each race in the labforce percentage. Black people working percentage is litter higher than white people.

```{r}
ggplot(data, aes(fct_infreq(LABFORCE), group = RACE)) + 
  geom_bar(aes(y = ..prop.., fill = LABFORCE), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  facet_wrap(~RACE) +
  xlab("Labforce") + 
  ylab("Percentage in Each Subgroup") +
  ggtitle("Labforce Percentage for each Race") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

2. Occupation Score for each Race: Although there is little difference between each race in working percentage, there is significant difference between occupation score. It shows white people have significantly higher salary than black people and other races. And the variance in salary for white people is lees than other races. Therefore, it means white people salary is high and centered.

```{r}
ggplot(data, aes(x=RACE, y=OCCSCORE)) +
  geom_boxplot() +
  coord_flip() + 
  ggtitle("Occupation Score for each Race") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

3. Occupation category for each Race: Black people tend to work as operatives, service workers and laborers. White people tend to work as craftsmen, clerical and kindred, managers and sales workers.

```{r}
ggplot(data) +
  geom_bar(aes(x=fct_infreq(OCC1950CAT), fill=RACE), stat = "count") +
  xlab("Occupation Category") +
  ylab("Count of Records") + 
  ggtitle("Occupation Category for each Race") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

4. Industry category for each race: Black people tend to work in personal services industry.

```{r}
ggplot(data) +
  geom_bar(aes(x=fct_infreq(IND1950CAT), fill=RACE), stat = "count") +
  xlab("Industry Category") +
  ylab("Count of Records") + 
  ggtitle("Industry Category for each Race") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

## Education and Employment

1. Labor force for each gender and literate: As shown in the figure, there is little difference between those who are literate and who are illiterate in male workers. However, female who are literate are more likely in the labor force.

```{r}
ggplot(data, aes(LABFORCE, group = LIT)) + 
  geom_bar(aes(y = ..prop.., fill = LABFORCE), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  facet_wrap(~SEX+LIT) +
  coord_flip() +
  xlab("Labforce") + 
  ylab("Percentage in Each Subgroup") +
  ggtitle("Labforce Percentage for each Gender and Literate") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

2. Labor force for each race and literate: From the following figure, the people who are literate tend to be in the labor force. Especially for white people, white people who are literate significantly tend to work in the labor force.

```{r}
ggplot(data, aes(LABFORCE, group = LIT)) + 
  geom_bar(aes(y = ..prop.., fill = LABFORCE), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  facet_wrap(~RACE+LIT) +
  coord_flip() +
  xlab("Labforce") + 
  ylab("Percentage in Each Subgroup") +
  ggtitle("Labforce Percentage for each Race and Literate") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

3. Occupation score for each race and literate: From the following figure, we conclude that people who are literate tend to have higher salary than those who are not literate. Especially for female workers, literate female workers median salary are significantly highers than illiterate female workers.

```{r}
ggplot(data, aes(x=LIT, y=OCCSCORE)) +
  geom_boxplot() +
  facet_wrap(~SEX) +
  ggtitle("Occupation Score for each Race and Literate") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

4. Occupation score for each race and literate: The figure shows literate people tend to have higher salary. Especially for black people, the median salary for literate black workers are higher than illiterate black workers.

```{r}
ggplot(data, aes(x=LIT, y=OCCSCORE)) +
  geom_boxplot() +
  facet_wrap(~RACE) +
  ggtitle("Occupation Score for each Race and Literate") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

5. Occupation category and literate level: The following figure shows illiterate people tend to work as operatives, craftsmen and laborers, and literate people tend to work as clerical and kindred, sales workers, and professional.

```{r}
ggplot(data) +
  geom_bar(aes(x=fct_infreq(OCC1950CAT), fill=LIT), stat = "count") +
  xlab("Occupation Category") +
  ylab("Count of Records") + 
  ggtitle("Occupation Category and Literate Level") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

6. Industry category and literate level: illiterate people tend to work in retail, manufacuring-nondurable goods, and personal services area. And literate people can work in industries like finance, business, public administrationm, etc.

```{r}
ggplot(data) +
  geom_bar(aes(x=fct_infreq(IND1950CAT), fill=LIT), stat = "count") +
  xlab("Industry Category") +
  ylab("Count of Records") + 
  ggtitle("Industry Category and Literate Level") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 
```

## Time series and Employment

1. Labor force for races and genders from 1850 to 1940: From the following figure, we could conclude that from 1850 to 1940, more and more female white people enter the labor force. The male workers start to enter the labore force from 1850 to 1920, and then some of them quit the labor force in 1930s and 1940s.

```{r}
ggplot(data, aes(x=YEAR, group = LABFORCE, fill=LABFORCE)) + 
  geom_bar(aes(y = ..prop.., fill = LABFORCE), stat="count", position = "fill") + 
  scale_y_continuous(labels=scales::percent) +
  facet_wrap(~SEX+RACE) +
  coord_flip() +
  scale_x_reverse(breaks = seq(1850, 1940, by = 10)) +
  ggtitle("Labor Force for Races and Genders from 1850 to 1940") +
  theme(legend.position="bottom", plot.title = element_text(hjust = 0.5))
```

2. Occupation score for races and genders from 1850 to 1940: As shown in the figure, from 1850 to 1940, the female occupation score increase, especially for white people. And the black male workers salary increase a little. The salary for white male is steady acoross different years.

```{r}
time_series <- data %>%
  group_by(YEAR, SEX, RACE) %>%
  summarise(
    OCCSCORE = mean(OCCSCORE, na.rm = TRUE)
  )
ggplot(time_series, aes(x=YEAR, y=OCCSCORE)) +
  geom_line(aes(color=RACE)) +
  scale_x_continuous(breaks = seq(1850, 1940, by = 10)) +
  theme(legend.position="bottom") + 
  facet_wrap(~SEX) +
  ggtitle("Occupation Score for Races and Genders from 1850 to 1940") +
  theme(legend.position="bottom", plot.title = element_text(hjust = 0.5))
```

## Conclusion

1. The minority working conditions are worse than the majority. Female workers who are in the labor force percentage is significantly less than the male workers, and the salary for female workers is less than male workers. There is little difference between difference races in the labor force. However, the salary for the male workers are less than white people.
2. There are significant difference in industry categories and types of work between male workers and female workers. Female workers tend to work in the industry like personal services as services workers, while male workers tend to work in construction, transportation, etc. as craftsman, laborer, etc. And the black people tend to work in personal services as services workers, while white people tend to work in officials, finance, etc. as managers or professionals.
3. Education levels significantly influence the working conditions. Literate people tend to have higher salary, and work at officials, business, finance area as professionals or managers. Illiterate people have lower salary, and work at construction, manufacturing as workers and craftsman.
4. From 1850 to 1940, minority people working conditions improve a lot. The salary and percentage of people in the labor force increase for female workers and black workers.

<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component



<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion


<!--chapter:end:07-conclusion.Rmd-->

